\documentclass[10pt]{extarticle}
\usepackage[margin = 1in]{geometry}
\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx, fancyhdr}
\usepackage{amsfonts, amssymb}
\usepackage{amsmath, amsthm, lipsum, mathtools}
\usepackage{xparse, tikz, multicol, enumitem, pgfplots}
\usepackage{algorithm, algpseudocode, hyperref}
\usepackage{stickstootext}
\usepackage[stickstoo,vvarbb]{newtxmath}
\usepackage[framemethod = TikZ]{mdframed}
\usetikzlibrary{shadows, shapes.geometric, positioning, arrows.meta}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{decorations.markings}
\pgfplotsset{compat=newest}

\hypersetup{
    colorlinks=true,
    linkcolor=ProcessBlue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={},
    }

\tikzset{
    lattice edge/.style = {draw, -},
}

\tikzset{
    u/.style = {lattice edge, to path={(\tikztostart) -- (\tikztotarget.south)\tikztonodes}},
    uu/.style = {lattice edge, to path={(\tikztostart.north) -- (\tikztotarget)\tikztonodes}},
}

\urlstyle{same}

\pagestyle{fancy}
\fancyhead{}
\fancyhead[RO]{\leftmark}
\fancyhead[LO]{\rightmark}
\setlength{\headheight}{14pt}
\addtolength{\topmargin}{-0.5pt}

\input{Misc Environments/breakable algorithm.tex}
\input{Base Tex Files/commands.tex}
\input{Base Tex Files/environments.tex}

% Extra Commands

\title{Solutions to Dummit \& Foote's Abstract Algebra $3^{\text{rd}}$ Edition}
\author{blanket}
\date{October $27^{\text{th}}$, 2025}

\setlength{\jot}{1pt}
\setlength{\parindent}{0.5cm}

\begin{document}


\maketitle

\tableofcontents

\newpage

\section*{Preface}
\addcontentsline{toc}{section}{\protect\numberline{}Preface}

This is a collection of solutions to the exercises in Dummit and Foote's \textit{Abstract Algebra}, $3^{\text{rd}}$ edition. These solutions were written by me as I worked through the book, and are intended to serve as a reference for myself and others who are studying abstract algebra. I have made every effort to ensure the correctness of these solutions, but I cannot guarantee that they are free of errors. If you find any mistakes or have suggestions for improvement, please feel free to contact me.

I've attempted to format the solutions in a clear and consistent manner, using LaTeX for typesetting. Each chapter's exercises are included in separate files for better organization if viewing on GitHub. My GitHub also contains my \verb|commands.tex| and \verb|environments.tex| files that I use for formatting and defining custom commands/environments in \LaTeX, which may prove useful to others in their own projects and notes. Moreover, solutions to exercises only utilize techniques that are available to an advanced undergraduate student or beginning graduate student, in line with the intended audience of the textbook. As an aside, I've graduated from a bachelor's program in mathematics but have not pursued graduate studies, so my background is primarily at the undergraduate level.

Some suggestions I've been given to improve the guide are citing particularly important exercises that will prove useful in reading subsequent chapters, and providing writeups/discussions for why I choose a particular method of solution (such as functions/homomorphisms that may appear out of nowhere). While I have not yet implemented these suggestions, I plan to do so in the future as I continue to expand and refine the guide with a current goal being to finish all exercises first, then go back and add these additional details and comments.

Many thanks to the authors, David S. Dummit and Richard M. Foote, for writing such an excellent textbook that has been a valuable resource for my recreational studies in abstract algebra. I've received requests on Reddit for individuals to assist me in completing this project, but at this time I prefer to work on it independently. However, I welcome feedback and suggestions from anyone who is interested in contributing to the project in the future.
\include{exercises/chap0ex}
\include{exercises/chap1ex}
\include{exercises/chap2ex}
\include{exercises/chap3ex}
\include{exercises/chap4ex}

\end{document}